Supplemental Material for ``Assessing United States county-level exposure for research on tropical cyclones and human health"

### 2019-07-29 writing/ehp_methods.tex
Add some on wind radii and Ida

To validate these modeled county-level wind speed estimates, we compared them
to county-level maximum sustained surface wind speeds based on the wind radii
estimates in \ac{HURDAT2}, which are based on \ldots. The wind radii values
have been included in \ac{HURDAT2} since 2004 [?] and estimate the distance
from the storm's center to the furthest point with winds at or above winds of a
certain speed in each four quadrants of the storm. They give estimates of these
maximum radii for three thresholds of maximum surface wind speeds: 64,~50,
and~34 \si{\knot}.  They therefore allow for the classification of counties
into four categories of maximum sustained wind speeds: $<$34~\si{\knot};
34\,--\,49.9~\si{\knot}; 50\,--\,63.9~\si{\knot}; $\ge$64~\si{\knot}. We
interpolated this data to~15-\si{\minute} increments and classified a county as
exposed to winds in a given wind speed category if its population mean center
was within~85\% of the maximum radius for that wind speed in its quadrant of
the storm [clarify why 85\%]. We compared these wind radii-based estimates with
the modeled wind speed estimates for the study storms since 2004 for which at
least one study county had a sustained wind of~$\ge$34 \si{\knot} (based on the
\ac{HURDAT2} wind radii). For each of these storms, we calculated the percent
of study counties that were classified in the same wind speed category.

>>>>

To validate these modeled county-level wind speed estimates, we compared them
to county-level maximum sustained surface wind speeds estimated based on the
wind radii estimates in \ac{HURDAT2}, which have been included in \ac{HURDAT2}
since 2004 [ref]. Reanalysis values for these wind radii estimate the maximum
extent from the storm's center to winds at or above winds of a certain speed in
each four quadrants of the storm, based on data from satellites and, where
available, ground-based and aircraft reconnaissance data [ref].  Wind radii are
estimated for three thresholds of maximum surface wind speeds: 64,~50, and~34
\si{\knot}.  They therefore allow for the classification of counties into four
categories of maximum sustained wind speeds: $<$34~\si{\knot};
34\,--\,49.9~\si{\knot}; 50\,--\,63.9~\si{\knot}; $\ge$64~\si{\knot}. We
interpolated this data to~15-\si{\minute} increments and classified a county as
exposed to winds in a given wind speed category if its population mean center
was within~85\% of the maximum radius for that wind speed in its quadrant of
the storm, as previous research has found  ... [ref]. We compared these wind
radii-based estimates with the modeled wind speed estimates for the study
storms since 2004 for which at least one study county had a sustained wind
of~$\ge$34 \si{\knot} (based on the \ac{HURDAT2} wind radii). For each of these
storms, we calculated the percent of study counties that were classified in the
same wind speed category.

### 2019-07-29 writing/ehp_results.tex
Add some on wind radii and Ida

For flood data, we compared flood status, based on event listings in the NOAA
Storm Events database, to streamflow measurements at \ac{US} Geological Survey
gages within nine sample counties (Figure~\ref{fig:floodcomparison}).  Across
the sampled counties, streamgage data generally indicated higher discharge in
on dates identified as flood events, particularly in counties with more
reporting stream gages (e.g., Harris County, TX, with~10 reporting stream
gages; Figure~\ref{fig:floodcomparison}). There were some cases, however, where
the two flooding data sources were somewhat inconsistent.  For example, there
were one or two tropical cyclones in several of the counties (Mobile County,
AL, Escambia County, FL, and Montgomery County, PA) which were classified as
flood events based on NOAA storm event listings but for which the total
discharge across county streamflow gages was below the 25\textsuperscript{th}
percentile of measurements taken during storms that did not have a flood event
listing. For tropical cyclones that did not have an associated flood listing in
the NOAA Storm Events database, most had lower total streamflow discharge
across county gages compared to storms with flood event listings, and in all
cases but one all gages were below the flooding threshold.  The exception was
one storm in Fulton County, GA (Ida in 2009). [More on which storm and why---I
think there was already a flood event that had started earlier, so it missed
the cutoff for the startdate being within the range.]

>>>>

For flood data, we compared flood status, based on event listings in the NOAA
Storm Events database, to streamflow measurements at \ac{US} Geological Survey
gages within nine sample counties (Figure~\ref{fig:floodcomparison}).  Across
the sampled counties, streamgage data generally indicated higher discharge in
on dates identified as flood events, particularly in counties with more
reporting stream gages (e.g., Harris County, TX, with~10 reporting stream
gages; Figure~\ref{fig:floodcomparison}). There were some cases, however, where
the two flooding data sources were somewhat inconsistent.  For example, there
were one or two tropical cyclones in several of the counties (Mobile County,
AL, Escambia County, FL, and Montgomery County, PA) which were classified as
flood events based on NOAA storm event listings but for which the total
discharge across county streamflow gages was below the 25\textsuperscript{th}
percentile of measurements taken during storms that did not have a flood event
listing. For tropical cyclones that did not have an associated flood listing in
the NOAA Storm Events database, most had lower total streamflow discharge
across county gages compared to storms with flood event listings, and in all
cases but one all gages were below the flooding threshold.  The exception was
one storm in Fulton County, GA (Ida in 2009). This storm occurred in November
2009, following a month with historic rainfall and flooding in Georgia [ref].
Therefore, in this case the flooding associated with Ida was incorporated into
an ongoing flood event listing, with a start date well before the five-day
window used to temporally match storm event listings with tropical cyclone
tracks used here.
